Importance sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body

### 2019-04-15 draft_p2.tex
Minor modifications in part IV.B.2 results GKTL

We now discuss the application of the \ac{gktl} algorithm to the flow dynamics introduced in section~\ref{sec:test_flow}.
The objective is to sample trajectories displaying extreme fluctuations of the \textit{time-averaged} drag~\eqref{eq:def_time_averaged_drag}.
Consistently with section~\ref{sec:time_avg}, we choose the averaging window $T=10\tau_c$.

% Choice of parameters and perturbation
The computational cost $C_{gktl}$ of a \ac{gktl} experiment is determined by both the duration of the trajectories $T_a$ and the number of trajectories $N_c$: $C_{gktl} = N_c \times T_a$.
In the following series of experiments we set $T_a = T = 10\tau_c$.
The sampling from the \ac{gktl} algorithm will be compared to the sampling from the control run, \textit{i.e.} a direct simulation of the flow over $T_{tot}=10^6\tau_c$.
We set $N_c=16384$, leading to $C_{gktl} \approx 1.6 \times 10^5 \tau_c$. (\textbf{Plus vraiment clair pour moi pourquoi je n'ai pas utilisé 10X plus de clones, pour matcher le coût numérique de la simulation de contrôle...})
Lastly, one must set the cloning period $\tau$.
On the one hand, too small a cloning period may result in loss of information: clones do not separate from their parent over a cloning period.
On the other hand, choosing $\tau \gg \tau_c$ may result in too few cloning stages to achieve importance sampling.
A rule of thumb is to set $\tau \approx \tau_c$.
In the following, we set $\tau = \tau_c /2$.

% Commentaire de la figure illustration_IS

We performed three experiments, corresponding to three different values of the parameter $k$.
In general the value of $k$ must be set empirically, unless a \textit{large deviation regime} $T_a \to \infty$ is verified~\cite{ref_ld_regime}.
In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem~\cite{GEref}.
% In the large time limit $T_a\to \infty$, the typical level of fluctuations $\mathbb{E}_{\mathbb{P}_k}[ F_{T_a} ]$ within the biased ensemble of trajectories
% can be related to the \textit{Scaled Cumulant Generating Function} $\lambda(k)$~\cite{ref}.

Figure~\ref{fig:illustration_IS} illustrates importance sampling achieved by means of the \ac{gktl} algorithm for the time-averaged drag $F_T$ with $T = T_a = 10\tau_c$.
For each of the three experiments, it displays the estimate of the biased \ac{pdf}
\begin{equation}
  \label{eq:estimate_biased_measure}
  \rho_k(F) = \mathbb{E}_{\set{P}_k}[\delta(F_{T_a}[\{\mathbf{x}\}_{0\leq t \leq T_a}]-F)]
  \approx \frac{1}{N_c}\sum_{j=1}^{N_c}\delta(F^{(j)}_{T_{a}}[\{\mathbf{x}\}_{0\leq t \leq T_a}] - F),
\end{equation}
computed over the set of $N=16384$ trajectories sampled by the algorithm.
Moreover, figure~\ref{fig:illustration_IS_drag} displays an estimate of the <<unbiased>> \ac{pdf} $\set{P}_0$ describing the time-averaged drag process.
Note that it corresponds to equation~\eqref{eq:estimate_biased_measure} with $k=0$.

@@

% Malgré les limitations, c'est toujours mieux
% qu'un échantillonnage direct
Unsurprisingly, sampling larger fluctuations requires a larger population size, and therefore a larger computational cost.
Even though, table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample trajectories corresponding to values of the time-averaged drag that could
not be sampled through a direct sampling with equal computational cost.
Table~\ref{tbl:nbTraj_tbl} displays the number of trajectories that correspond to a time-average $F_{T_a} \geq a$, for several values of $a$.
As a comparison, it also displays the average number of such fluctuations in a control timeseries of duration $C_{gktl}$, divided in $N_c = C_{gktl} / T_a$ events.
This number is computed based on the assumption that the \ac{pdf} describing the random variable $F_T$ with $T=10\tau_c$ is Gaussian.
This hypothesis is motivated by figure~\ref{fig:PDF_AVG}.
While this assumption does not hold for $F_T \geq 4\sigma_T$, it may still provide numbers relevant to the comparison with the algorithm.
\begin{figure}
  \centering
  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
  \caption{PDF of average drag over $10\tau_c$}
  \label{fig:PDF_AVG}
\end{figure}

>>>>

We now discuss the application of the \ac{gktl} algorithm to the flow dynamics introduced in section~\ref{sec:test_flow}.
The objective is to sample trajectories displaying extreme fluctuations of the \textit{time-averaged} drag~\eqref{eq:def_time_averaged_drag}.
Consistently with section~\ref{sec:time_avg}, we set the averaging window as $T=10\tau_c$.

% Choice of parameters and perturbation
The computational cost $C_{gktl}$ of a \ac{gktl} experiment is determined by both the duration of the trajectories $T_a$ and the number of trajectories $N_c$: $C_{gktl} = N_c \times T_a$.
In the following we set $T_a = T = 10\tau_c$.
We set $N_c=16384$, leading to $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
Lastly, one must set the cloning period $\tau$.
On the one hand, too small a cloning period may result in loss of information: clones do not separate from their parent over a cloning period.
On the other hand, choosing $\tau \gg \tau_c$ may result in too few cloning stages to achieve importance sampling.
As a consequence, a rule of thumb is to set $\tau \approx \tau_c$.
In the following, $\tau = \tau_c /2$.

We performed three experiments, corresponding to three different values of the parameter $k$.
In general the value of $k$ must be set empirically, unless a \textit{large deviation regime} $T_a \to \infty$ is verified~\cite{ref_ld_regime}.
In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem~\cite{GEref}.
Each experiment was run using $512$ Intel processors, each processor being assigned the computation of $N_c / 512 = 32$ clones.
The walltime for one experiment was roughly two days.

% Commentaire de la figure IS_GKTL
Figure~\ref{fig:illustration_IS} illustrates importance sampling achieved by means of the \ac{gktl} algorithm for the time-averaged drag $F_T$ with $T = T_a = 10\tau_c$.
For each of the three experiments, it displays the estimate of the biased \ac{pdf}
\begin{equation}
  \label{eq:estimate_biased_measure}
  \rho_k(F) = \mathbb{E}_{\set{P}_k}[\delta(F_{T_a}[\{\mathbf{x}\}_{0\leq t \leq T_a}]-F)]
  \approx \frac{1}{N_c}\sum_{j=1}^{N_c}\delta(F^{(j)}_{T_{a}}[\{\mathbf{x}\}_{0\leq t \leq T_a}] - F),
\end{equation}
computed over the set of $N=16384$ trajectories sampled by the algorithm.
Moreover, figure~\ref{fig:illustration_IS_drag} displays an estimate of the <<unbiased>> \ac{pdf} $\set{P}_0$ describing the time-averaged drag process.
Note that it corresponds to equation~\eqref{eq:estimate_biased_measure} with $k=0$.

@@

% Malgré les limitations, c'est toujours mieux
% qu'un échantillonnage direct
Unsurprisingly, sampling larger fluctuations requires a larger population size, and therefore a larger computational cost.
Nevertheless, table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample trajectories corresponding to values of the time-averaged drag that could
not be sampled through a direct sampling with equal computational cost.
Table~\ref{tbl:nbTraj_tbl} displays the number of trajectories that correspond to a time-average $F_{T_a} \geq a$, for several values of $a$.
As a comparison, it also displays the average number of such fluctuations in a control timeseries of duration $C_{gktl}$, divided in $N_c = C_{gktl} / T_a$ events.
This number is computed based on the assumption that the \ac{pdf} describing the random variable $F_T$ with $T=10\tau_c$ is Gaussian.
This hypothesis is motivated by figure~\ref{fig:PDF_AVG}.
While this assumption does not hold for $F_T \geq 4\sigma_T$, it may still provide numbers relevant to the comparison with the algorithm.
\begin{figure}
  \centering
  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
  \caption{PDF of average drag over $10\tau_c$}
  \label{fig:PDF_AVG}
\end{figure}
